---
title: "Human Midbrain Embryonic Dataset"
author: "Nathan Skene"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Human Midbrain Embryonic Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc} 
---

<!-- Readme.md is generated from Readme.Rmd. Please edit that file -->

```{r style, echo=FALSE, results='asis', message=FALSE}
BiocStyle::markdown()
knitr::opts_chunk$set(tidy         = FALSE,
                      warning      = FALSE,
                      message      = FALSE)
```

## Installation

```
install.packages("devtools")
library(devtools)
install_github("nathanskene/CTD.human.midbrain.embryonic")
```

## Human Midbrain Embryonic Dataset

This package provides easy access to the human midbrain embryonic dataset, which is associated with the following paper:

[La Manno G, Gyllborg D, Codeluppi S, Nishimura K et al. Molecular Diversity of Midbrain Development in Mouse, Human, and Stem Cells. Cell 2016 Oct 6;167(2):566-580.e19. PMID: 27716510](https://www.cell.com/cell/fulltext/S0092-8674(16)31309-5)

The raw data is available along with annotations:

```{r ,eval=FALSE}
library(CTD.human.midbrain.embryonic)
data(raw_sc_expression_data)
raw_sc_expression_data$exp[1:5,1:5]
raw_sc_expression_data$annot[["level2class"]][1:10]
raw_sc_expression_data$annot[["level1class"]][1:10]
```

Should you wish to regenerate the data from scratch then the code for doing so is available in `process_raw_sc_data()`. I suggest forking this project, modifying the code for this function in there.

The processed cell type data (ctd) for use in MAGMA_Celltyping and EWCE is also available using:

```{r }
library(CTD.human.midbrain.embryonic)
data(ctd)
```

## Preparing cell type data file

```{r, eval=FALSE}
library(EWCE)
raw_sc_expression_data$exp_CORRECTED = fix.bad.hgnc.symbols(raw_sc_expression_data$exp)
raw_sc_expression_data$exp_CORRECTED_DROPPED = drop.uninformative.genes(exp=raw_sc_expression_data$exp_CORRECTED, level2annot = raw_sc_expression_data$annot$level2class)
fNames = generate.celltype.data(exp=raw_sc_expression_data$exp_CORRECTED_DROPPED,raw_sc_expression_data$annot,"HumanEmbyronicMidbrain")
```

We can check the results are as expected:

```{r, eval=FALSE}
load(fNames[1])
print(ctd[[1]]$specificity[1:5,1:5])
print(sort(ctd[[1]]$specificity[,"Embryonic Dopaminergic Neuron"],decreasing = TRUE)[1:20])
```
